<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{ISMB Twitter Analysis}
%\VignetteDepends{ggplot2}
-->

# ISMB Twitter Analysis

The goal of this work is to analyze two different collections of archived tweets from the ISMB conferences from 2012 and 2014, and compare and contrast them. The collections of tweets are from Neil Saunders ([his initial analysis](http://nsaunders.wordpress.com/2012/08/16/twitter-coverage-of-the-ismb-2012-meeting-some-statistics/) and [data](https://github.com/neilfws/Twitter/tree/master/ismb/data)) and Stephen Turner ([initial analysis](http://gettinggeneticsdone.blogspot.com/2013/05/automated-analysis-tweets-bioinformatics-twitterchive.html) and [data](https://github.com/stephenturner/twitterchive)).

Before we start to use this data, we will do some munging on it so that both data-sets are comparable and we can apply the same functions to them for comparison.

```{r datamunging}
baseLoc <- system.file(package="ismbTweetAnalysis")
extPath <- file.path(baseLoc, "extdata")
```

```{r data2012, eval=FALSE}
load(file.path(extPath, "ismb2012.RData"))
ismb <- ismb[, c("text", "created", "id", "screenName")]
ismb$hashSearch <- "ismb"
save(ismb, file="data/ismb2012.RData")
```

This is especially important for the 2014 data, because it is not in `rdata` format, and there were actually three different hashtags searched for. 

```{r data2014, eval=FALSE}
ismb <- readTweetData(file.path(extPath, "ismb.txt"), "ismb")
ismb2014 <- readTweetData(file.path(extPath, "ismb2014.txt"), "ismb2014")
ismb14 <- readTweetData(file.path(extPath, "ismb14.txt"), "ismb14")
ismb <- rbind(ismb, ismb14, ismb2014)
save(ismb, file="data/ismb2014.RData")
```

# Analysis

Lets do some simple things in each case. For example, we can look at the distribution of tweets by time, who had the most `retweets`, who was the most prolific tweeter (as a percentage of total tweets), etc, and even for changes between the two years. Note that this is not the most comprehensive analysis we could do, because this is mostly an example analysis of how to do an analysis as a package vignette, but I need to do something, right?

## 2012 

```{r load2012}
data(ismb2012)
head(ismb)
```

```{r visualize2012}
library(ggplot2)
ggplot(ismb, aes(x=created)) + geom_bar()
```

